---
title: "Sales Data Analysis Report"
author: "Celumusa Fakudze"
date: "2024-05-12"
output: html_document
---

## Data Importing into R-studio Environment
The Auto Sales Data was downloaded from the Kaggle website.

```{r}
library(readr)
Auto_Sales_data <- read_csv("C:/Users/Fakudze/Desktop/Auto Sales data.csv")
View(Auto_Sales_data)
```

## Loading Libraries for data pre-procesing and analysis

```{r}
library(rmarkdown)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(summarytools)
library(lubridate)
```

## **Data Cleaning and Sampling**

## Checking if there are any missing values in the dataset

```{r}
sum(is.na(Auto_Sales_data))
```

## Data Sampling
Notes: Variables such as the ORDERNUMBER, QUANTITYORDERED, PRICEEACH, ORDERDATE, PRODUCTCODE, PRODUCTLINE, STATUS, CUSTOMERNAME, CITY, COUNTRY, DEALSIZE and SALES are selected for the analysis.The new data sample **data** is produced.

```{r}
data <- Auto_Sales_data %>% 
           select(ORDERNUMBER, QUANTITYORDERED, PRICEEACH, ORDERDATE, PRODUCTCODE, PRODUCTLINE, STATUS,
                  CUSTOMERNAME, CITY, COUNTRY, DEALSIZE, SALES)
```

## Examining data types types
Checking the data types of the variables

```{r}
str(data)
```

## Transforming the ORDERDATE variable from character to date format


```{r}
data$new_date <- dmy(data$ORDERDATE)
```


## Data modelling
Create the year, day_of_month, month and weekdays variables


```{r}
x <- as.Date(data$new_date)
year <- as.factor(format(x, "%Y"))
day_of_month <- as.numeric(format(x, "%d"))
month <- lubridate::month(data$new_date, label = T, abbr = F,)
weekdays <- lubridate::wday(data$new_date, label = T, abbr = F)
```


## Creating new dataset using the cbind fuction with the year, day_of_month, month and weekdays variables 
The new dataset is called **data_new**


```{r}
data_new <- cbind(data, year, month,weekdays, day_of_month)
View(data_new)
```


## Removing the old date column ORDERDATE from the new dataset


```{r}
data_new <- data_new %>% 
           select(-ORDERDATE)
```


## Transforming PRODUCTLINE, STATUS, DEALSIZE into factor variables


```{r}
data_new$PRODUCTLINE <- as.factor(data_new$PRODUCTLINE)
data_new$STATUS <- as.factor(data_new$STATUS)
data_new$DEALSIZE <- as.factor(data_new$DEALSIZE)
str(data_new)
```

## **Exploratory Data Analysis**
Examining the the measures of central tendency and sales perfomancies across categorical variables


## Calculating descriptive statistics for numeric variables to see their distribution


```{r}
numerics <- data_new %>% 
           select(SALES, QUANTITYORDERED, PRICEEACH)

numerics %>% 
           descr(order = "preserve",
                 stats = c("mean", "sd", "min", "q1", "med",
                           "q3", "max"),
                 round.digits = 3)
```


### SALES trends on categorical variables


```{r}
categorical_features <- data_new %>% 
           select(STATUS, DEALSIZE, PRODUCTLINE, COUNTRY,
                  DEALSIZE, SALES)

cat_long_features <- categorical_features %>%
           pivot_longer(!SALES, names_to = "features", values_to = "values") %>% 
           group_by(features) %>% 
           mutate(values = factor(values))

cat_long_features %>% 
           ggplot()+
           geom_bar(aes(x = values, fill = features), alpha = 0.7,
                    show.legend = F)+
           facet_wrap(~features, scales = 'free')+
           paletteer::scale_fill_paletteer_d("ggthemr::solarized")+
           theme(panel.grid = element_blank(),
                 axis.text.x = element_text(angle = 90))
```


## Time Series analysis

### Sales trend by days of a week


```{r}
data_new %>% 
           ggplot(aes(weekdays, SALES))+
           geom_boxplot(color = "red")

```


### Sales trend by month of year


```{r}
data_new %>% 
           ggplot(aes(x = month,  y = SALES))+
           geom_boxplot(color = "blue")
```

### Sales trnd by day of month


```{r}
plot(data_new$SALES, xlab = "Day of the month", ylab = "Sales", lwd = 2, type = "l", xlim = c(1, 31), ylim = c(500, 15000), main = "Sales Gragh by days of the month")
```

## Customername with highest sales


```{r}
data_new %>% 
           select(CUSTOMERNAME, SALES, CITY, COUNTRY, ) %>% 
           group_by(CUSTOMERNAME) %>% 
           arrange(desc(SALES))
```
